---
title: "acceptTime_analyze"
author: "Ian Sturgill"
date: "2023-11-29"
output: html_document
---
# acceptTime_analyze

## Purpose
This R markdown document documents the process of analyzing data
previously extracted from PLOS article XML files.

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

First we'll load the required libraries and data.

```{r, echo = TRUE, eval = FALSE}
library(docstring)
library(ggplot2)
library(ggdist)

dat <- read.delim("data/2023-11-29_plostime_data.txt")
```

Now we need a function to subset the dataframe based on a specific journal
and/or publication date cutoff:

```{r, echo = TRUE}
dat_subset <- function(dat, journal = NULL, cutoff = 180) {
  #' Dataframe subset
  #' 
  #' Returns a dataframe subset based on journal and/or publication date cutoff
  #' @param dat R dataframe object
  #' @param journal PLOS journal abbreviation string, e.g. pbio
  #' @param cutoff Publication date cutoff, in days from current date
  
  if (!is.null(journal)) {
    if (!(journal %in% unique(dat$journal))) {
      stop(sprintf("%s not in list of journals", journal))
    }
  }
  
  # Compute time in days between today and publication date
  today <- Sys.Date()
  time_diff <- as.integer(difftime(today, dat$epub_date))
  
  if (is.null(journal)) {
    dat_sub <- dat[time_diff < cutoff, ]
  }
  
  else {
    dat_sub <- dat[dat$journal == journal & time_diff < cutoff, ]
  }
  
  return(dat_sub)
}
```

Let's use the default function values to subset to articles published in any
PLOS journal in the last 180 days, or about 6 months.

```{r, echo = TRUE}
dat_6mo <- dat_subset(dat)
```

Now we'll create raincloud plots to show the distributions of acceptance times.

```{r, echo = TRUE, warning = FALSE}
# Manually retrieved the colors used in the Python notebook data visualization

sns_colors <- c('#db5f57',
 '#dba157',
 '#d3db57',
 '#91db57',
 '#57db5f',
 '#57dba1',
 '#57d3db',
 '#5791db',
 '#5f57db',
 '#a157db',
 '#db57d3',
 '#db5791')

names(sns_colors) <- c("pbio",
                       "pcbi",
                       "pclm",
                       "pdig",
                       "pgen",
                       "pgph",
                       "pmed",
                       "pntd",
                       "pone",
                       "ppat",
                       "pstr",
                       "pwat")

# Raincloud plots from geom_boxplot and stat_halfeye
# Reorder journals from highest to lowest median acceptance times
ggplot(dat_6mo,
       aes(x = reorder(journal, accept_time, 
                       function(x) {median(x, na.rm = TRUE)}, 
                       decreasing = TRUE), 
           y = accept_time,
           fill = journal)) +
  geom_boxplot(outlier.shape = NA,
               width = 0.12,
               show.legend = FALSE) +
  stat_halfeye(adjust = 0.5,
               width = 0.6,
               .width = 0,
               justification = -0.2,
               point_color = NA,
               show.legend = FALSE) +
  theme_classic() +
  scale_fill_manual(values = sns_colors) +
  ggtitle("PLOS time to acceptance of articles published in last 180 days") +
  xlab("Journal") +
  ylab("Time to acceptance (days)") +
  theme(plot.title = element_text(hjust = 0.5))
```


# End of R markdown document